---
title: "Assignment1"
output:
  html_document:
    df_print: paged
date: "2023-07-31"
---

These exercises require you to generate plots of various kinds. 


## Part I -- geoms and aesthetics

These first few exercises will run through some of the simple principles of
creating a ggplot2 object, assigning aesthetics mappings and geoms.

1. Read in the cleaned patients dataset, `patient-data-cleaned.txt`, into a
   new object called `patients`.

```{r message = FALSE, warning = FALSE}
library(ggplot2)

patients = read.table("patient-data-cleaned.txt", header=T, sep="\t", fill=T)

head(patients)
```

### Scatterplots

2. Generate a scatterplot of BMI versus Weight using the patient dataset and add
   a colour scale based on the Height variable.

```{r}
s = ggplot(data = patients, mapping = aes(x = BMI, y = Weight)) +
  geom_point(mapping = aes(color = Height))

s
```

3. Using an additional geom, add an extra layer of a fit line to the previous plot.(use geom_smooth() )

```{r message = FALSE}
s + geom_smooth()
```


### Boxplots and Violin plots

4. Generate a boxplot of Score values comparing smokers and non-smokers.

```{r}
ggplot(data = patients, mapping = aes(x = Smokes, y = Score)) +
  geom_boxplot()
```

5. Split the previous boxplot into male and female groups with different colours.

```{r}
ggplot(data = patients, mapping = aes(x = Smokes, y = Score)) +
  geom_boxplot(mapping = aes(color = Sex))
```

### Histogram and Density plots

6. Generate a histogram of BMIs with each bar coloured blue, choosing a
    suitable bin width.

```{r}
ggplot(data = patients, mapping = aes(x = BMI)) +
  geom_histogram(binwidth = 0.5, mapping = aes(fill = "blue"))
# I do not why it outputs as red 
```

7. Instead of a histogram, generate a density plot of BMI

```{r}
ggplot(data = patients, mapping = aes(x = BMI)) +
  geom_density()
```