---
title: "Practice1"
author: "Baixi Guo"
date: "2023-09-03"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

```

# Section 1: Types of Plot
1. Scatter Plot
  + Purpose: Visualize relationship between two continous variables
  + Benefit: Identify patterns, correlations & outlier

2. Line Plot
  + Purpose: Display data point over time
  + Benefit: Observe trend over time

3. Bar Plot
  + Purpose: Compare categories points
  + Benefit: Frequency Count

4. Histogram Plot
  + Purpose: Visualize distribution and frequency of single continuous variable
  + Benefit: Understand distribution, central tendency and spread

5. Box Plot
  + Purpose: Display the distribution of a single continuous/categorical variable and its summary statistics
  + Benefit: Identify central tendency, spread, skewness and outliers

# Section 2: Plot Components
1. Dataset
2. Mapping (x,y) & Aesthetic
3. Layers
  + Geometric Object
  + Statistical Transformation
  + Position Adjustment
4. Scale
5. Coordinate System
6. Facet Specification
7. Annotation

# Section 3: Example For Components
```{r echo=FALSE}
# Setup
A <- c(2,1,4,9)
B <- c(3,2,5,10)
C <- c(4,1,15,80)
D <- c("a", "a", "b", "b")
simple <- data.frame(A,B,C,D)

# Simple Version
ggplot(data = simple, mapping=aes(x=A, y=C, shape=D)) +
  geom_point()

# Full Version
ggplot(data = simple, mapping=aes(x=A, y=C, shape=D)) +
  layer(geom = "point",
        stat = "identity",
        position = "identity") +
  scale_x_continuous() +
  scale_y_continuous() +
  coord_cartesian() +
  facet_null()
```