---
title: "Practice3"
author: "Baixi Guo"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 0 - Import Library
```{r}
library(dplyr)
library(tidyverse)
library(nycflights13)
```


## Section 1 - Visualize Dataset
```{r}
# View Dataset
head(flights)
```


## Section 2 - Use dplyr Function
### (Filter)
```{r}
df1 <- filter(flights, month==10)                                # Subset rows with one condition  
df2 <- filter(flights, month==10, carrier=="UA", origin=="EWR")  # Subset rows with multiple conditions
```
### (Arrange)
```{r}
df1 <- arrange(flights, dep_delay)          # Re-order rows in ascending order 
df2 <- arrange(flights, desc(dep_delay))    # Re-order rows in descending order
```
### (Select & Rename)
```{r}
df1 <- select(flights, year)                             # Subset cols with one condition
df2 <- select(flights, year, month, day, dep_time)       # Subset cols with multiple conditions
df3 <- select(flights, -year, -month, -day, -dep_time)   # Subset cols except multiple conditions

df4 <- select(flights, 2:5)                              # Subset cols starting and ending with indexes
df5 <- select(flights, month:sched_dep_time)             # Subset cols starting and ending with names

df10 <- rename(flights, YEAR=year, MONTH=month, DAY=day) # Subset cols to only to rename, but keep the same dataframe
```
### (Mutate & Transmute)
```{r}
# Create new col with defined formula
df1 <- mutate(flights, speed=distance/air_time*60)
# Create multiple new cols with defined formula
df2 <- mutate(flights, speed=distance/air_time*60, gain=arr_delay-dep_delay, acceleration=speed/air_time*60) 

# Create multiple new cols with defined formula and drop other cols
df3 <- transmute(flights, speed=distance/air_time*60, gain=arr_delay-dep_delay, acceleration=speed/air_time*60)
```
### (Summarize)
```{r}
# Summarize various statistics
df4 <- summarise(flights,med_dist=median(distance), 
                 mean_dep_delay=mean(dep_delay, na.rm=TRUE), 
                 mean_arr_delay=mean(arr_delay, na.rm=TRUE))
```


## Section 3 - Other Functions for dplyr Functions
### Other Function for Filter()
```{r}
# Relational Operators
# >, >=, <, <=, !=

# Logical Operators
# &, |, !

# Check for floating point equality
# x1 ~ x2
v1 = dplyr::near(0.055555, 0.055554)

# Check for between numbers
# a <= x | x <= b
v2 = dplyr::between(4:8, 3, 7)

# Check if a element exist in set
# x == a | x == b | x ==c ...
v3 = 3 %in% c(3,4,6,8)

# Check if NA value
# NA == NA
v4 = is.na(NA)
```
### Other Function for Select()
```{r}
df1 <- select(flights, dplyr::starts_with("arr"))               # Subset cols with "arrxxx" name
df2 <- select(flights, dplyr::ends_with("time"))                # Subset cols with "xxxtime" name
df3 <- select(flights, dplyr::contains("arr"))                  # Subset cols containing "arr"
df4 <- select(flights, dplyr::num_range("x", 1:3))              # Subset cols with name and number range
```
### Other Function for Mutate()
```{r}
# Arithmetic Operator: +,-,*,/,^
# Recycle shorter variables

# Modular Arithemtic Operator: %%, %/%
# Breaking apart integers (time into hours+minutes)

# Log Function: log(), log2(), log10()
# For data with multiplicative variance

# Cumulative Summary Function: cumsum(), cumprod(), cummax(), cummin()
# Use for cumulative statistics
r1 = cumsum(1:3)
r2 = cumprod(1:3)
r3 = cummax(1:3)
r4 = cummin(1:3)

# Relational Operators: >, >=, <, <=, !=
# Continuous variables --> Categorical variables

# Function: n()
# Can only be use in mutate(), filter() & summarize()
```
### Other Function for Summarize()
```{r}
# Measure of Location: mean(), median()

# Measure of Spread: sd(), var(), IQR()

# Measure of Rank: min(), max(), quantile()

# Count Functions: 
## dplyr::n(): number of observations
## sum(!is.na(x)): number of non-NA values
## dplyr::n_distinct(): number of unique values
## sum(x == a): number of occurrence of x that is value of a
## mean(x == a): proportion of x that is value of a

# group_by() & ungroup() Function
## group_by() used with summarize()
df1 = summarize(dplyr::group_by(flights, carrier), mean(arr_delay, na.rm=TRUE)) # Summarize mean by each group_by() category
## ungroup() ungroup a grouped dataset 
df2 = dplyr::ungroup(df1)
```


# Section 4 - Pipeline Operator
## Example 1
```{r}
df1 = filter(flights, month==10)

flights %>%
  filter(month==10) -> df2
```
## Example 2 & 3 Setup
```{r}
baz <- function(x) {
  y = " + baz"
  return (paste(x,y))
}

bar <- function(x) {
  y = " + bar"
  return (paste(x,y))
}

foo <- function(x){
  y = " + foo"
  return (paste(x,y))
}

stu <- function(x,y,z){
  return (paste(x,y,z))
}

x = "x"
```
## Example 2
```{r}
x1 = foo(bar(baz(x)))

x %>% baz() %>% bar() %>% foo() -> x2
```
## Example 3
```{r}
bazvar = baz(x)
barvar = bar(x)
foovar = foo(x)
x3 = stu(bazvar, barvar, foovar)

x %>% 
  {
    bazvar = baz(.)
    barvar = bar(.)
    foovar = foo(.)
    stu(x=bazvar, y=barvar, z=foovar)
  } -> x4
```